---
title: "Experiments for <i>Guided Visual Exploration of Relations in Data Sets</i>"
output: html_notebook
---

Experiments for manuscript 'Guided Visual Exploration of Relations in Data Sets' 
by Kai Puolamäki, Emilia Oikarinen, and Andreas Henelius

```{r}
rm(list = ls())
require(e1071)
require(RColorBrewer)
require(xtable)
require(fastICA)
source("../../src/R/corand.R")

sel_orange <- rgb(241 / 255, 163 / 255, 64 / 255)
sel_blue <- rgb(153 / 255, 142 / 255, 195 / 255)
sel_green <- rgb(0 / 255, 150 / 255, 115 / 255)
```

```{r global_options}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


# Toy example

```{r}
set.seed(42)
n <- 100
toydata <- matrix(NA,n,4,dimnames=list(NULL,LETTERS[1:4]))
a <- rnorm(n,mean=rep(c(-1,1),each=n/2),sd=0.25)
toydata[,"A"] <- rnorm(n,mean=a,sd=0.05)
toydata[,"B"] <- rnorm(n,mean=a,sd=0.05)
toydata[,"C"] <- rnorm(n,mean=toydata[,"A"],sd=0.2)
toydata[,"D"] <- rnorm(n,mean=toydata[,"B"],sd=0.2)
col <- rep(c("blue","red"),each=n/2)
t1 <- tiling(n,4)
t2 <- t1$copy()
t1$addtile(C=3:4) # H1
t2$addtile(C=3) # H2
t2$addtile(C=4) # H2
aa1 <- findproj2(t1$cov(toydata),t2$cov(toydata),k=1)
tt1 <- t1$copy()
tt2 <- t2$copy()
tt1$addtile(C=c(1,3)) # BG1
tt1$addtile(C=c(2,4)) # BG2
tt2$addtile(C=c(1,3)) # BG1
tt2$addtile(C=c(2,4)) # BG2
aa2 <- findproj2(tt1$cov(toydata),tt2$cov(toydata),k=1)
col <- "black" #rep(c("blue","red"),each=10)
p <- sample.int(2*n)
pdf("toy1.pdf")
pairs(toydata,col=col,cex=2)
dev.off()
pdf("toy2.pdf")
pairs(rbind(t1$permutedata(toydata),t2$permutedata(toydata))[p,],
      pch=rep(c(0,1),each=n)[p],
      col=rep(c(sel_blue,sel_orange),each=n)[p],cex=2)
dev.off()
pdf("toy3.pdf")
pairs(rbind(tt1$permutedata(toydata),tt2$permutedata(toydata))[p,],
      pch=rep(c(0,1),each=n)[p],
      col=rep(c(sel_blue,sel_orange),each=n)[p],cex=2)
dev.off()
aa1
aa2
```

The data has structure such that $A=B+small noise$, $C=A+large noise$, and $D=B+large noise$.

# Methods (Sec. 2.2)

Verify that the covariance approximation (Thm. 2) is numerically correct.

```{r}
set.seed(42)
## verify covariance theorem (Thm. 2)
vercov <- function(nmin=10000) {
  t0 <- tiling(13,7)
  data <- matrix(rnorm(13*7),13,7)
  for(i in 1:3) t0$addtile(sample.int(13,5),sample.int(7,3))
  cov_analytic <- t0$cov(data)
  cov_sample <- cov(t0$permutedata(data,nmin=nmin))
  max(abs(range(cov_analytic-cov_sample)))
}
for(nmin in c(10000,50000,250000,1000000)) {
  cat(sprintf("%d %g\n",nmin,vercov(nmin)))
}
```

# Stability and scalability (Sec. 3.1)

## Load the data
Set the random seed, load the data and select the appropriate variables.

```{r}


data_orig <- readRDS("../webdemo/webdemo/data/socio_economics_germany_full.rds")
data      <- data_orig[, 14:45]
data      <- scale(as.matrix(data))

data_real    <- data
data_factors <- data_orig[,3:5]

experiment <- function(x,f,k=3,nmin=1,e1=0,m1=0,n1=0,onlyy=FALSE) {
  ## Make messed up data
  # remove n1 rows
  y_rows <- sample.int(dim(x)[1],size=dim(x)[1]-n1)
  y_rows_rev <- rep(NA,dim(x)[1])
  y_rows_rev[y_rows] <- 1:length(y_rows)
  y <- x[y_rows,]
  # add random noise and rescale
  y <- scale(y+rnorm(length(y),sd=e1))
  # add m1 random columns by picking random columns from y and permuting them
  y <- cbind(y,apply(y[,sample.int(dim(x)[2],size=m1,replace=TRUE),drop=FALSE],2,sample))
  
  ## Generate a random tile with at least 2 columns
  randomtile <- function() {
    i <- sample.int(dim(f)[2],size=1) # Pick a random factor
    list(R=which(sample(levels(f[,i]),size=1)==f[,i]), # rows are from some factor
         C=sample.int(dim(x)[2],size=1+sample.int(dim(x)[2]-1,size=1))) # 2...m columns
  }
  
  ## Tiling etc. part
  time1 <- system.time({
    ## Create empty tilings
    if(!onlyy) ttx1 <- tiling(dim(x)[1],dim(x)[2])
    tty1 <- tiling(dim(y)[1],dim(y)[2])
  
    ## Add k random background tiles, with rows matching factors
    for(i in 1:k) {
      a <- randomtile()
      if(!onlyy) ttx1$addtile(a$R,a$C)
      R_y <- intersect(a$R,y_rows)
      if(length(R_y)>0) tty1$addtile(y_rows_rev[R_y],a$C)
    }
  
    if(!onlyy) ttx2 <- ttx1$copy()
    tty2 <- tty1$copy()
  
    ## Create two hypothesis tilings
    while(TRUE) {
      a <- randomtile()
      R_y <- intersect(a$R,y_rows)
      if(length(R_y)>0) {
        ## Make sure that the hypothesis tiling exists also in the messed up data
        if(!onlyy) {
          ttx1$addtile(a$R,a$C)
          for(i in a$C) ttx2$addtile(a$R,i)
        }
        tty1$addtile(y_rows_rev[R_y],a$C)
        for(i in a$C) tty2$addtile(y_rows_rev[R_y],i)
        break
      }
    }
    
  })
  
  ## Finding view part
  time2 <- system.time({
    if(!onlyy) {
      covx1 <- ttx1$cov(x)
      covx2 <- ttx2$cov(x)
      projx <- findproj2(covx1,covx2,k=1)
    }
    covy1 <- tty1$cov(y)
    covy2 <- tty2$cov(y)
    projy <- findproj2(covy1,covy2,k=1)
  })
  
  ## Rest is related more to debugging
  wy <- norm2(projy$w[1:dim(x)[2]])
  wx <- if(!onlyy) projx$w else wy
  wz <- norm2(rnorm(dim(x)[2]))
  
  
  if(onlyy) {
    l1 <- l2 <- l3 <- 1
  } else {
    l1 <- sum(wx*(covx1 %*% wx))/(sum(wx*(covx2 %*% wx)))
    l2 <- sum(wy*(covx1 %*% wy))/(sum(wy*(covx2 %*% wy)))
    l3 <- sum(wz*(covx1 %*% wz))/(sum(wz*(covx2 %*% wz)))
  }
  
  res <- c(l2/l1,l1,l2,l3,time1["elapsed"],time2["elapsed"])
  names(res) <- c("ratio","l1","l2","lrandom","time1","time2")
  res
}

```


## Robustness wrt. perturbation of data (Tab. 1a)

We try adding random noise and removing columns and see how it affects the results.

```{r}

res_pca <- matrix(NA,0,6,dimnames=list(NULL,c("nmin","e1","n1","m1","l1","l2")))

for(e1 in c(0,1,2,5,10)) {
  for(n1 in c(0,100,200)) {
    for(i in 1:49) {
      a <- experiment(data_real,data_factors,e1=e1,n1=n1)
      res_pca <- rbind(res_pca,matrix(c(0,e1,n1,0,a[2],a[3]),1,6))
    }
  }
}
saveRDS(res_pca,file="res_pca.rds")

```

Then plot the results.

```{r}
# Compute the mean relative error (percentage)
res_pca_mean <- aggregate((res_pca[,"l1"]-res_pca[,"l2"])/res_pca[,"l1"],
                          by=list(e=res_pca[,"e1"],n=res_pca[,"n1"],m=res_pca[,"m1"]),
                          mean)
res_pca_mean <- res_pca_mean[order(res_pca_mean[,1]),]
for(i in 1:dim(res_pca_mean)[1]) {
  cat(sprintf("$%g$ & $%d$ & $%.3f$ \\\\\n",
              res_pca_mean[i,1],
              res_pca_mean[i,2],
              res_pca_mean[i,4]))
}
```

## Scalability (Tab. 1b)

```{r}

createrandomdata <- function(n,m) matrix(rnorm(n*m),n,m)
createrandomfactors <- function(n,k=3) {
  a <- (n*k) %/% 2
  r <- as.data.frame(apply(matrix(c(rep(0,a),rep(1,n*k-a)),n,k,byrow=TRUE),2,sample))
  for(i in 1:k) r[,i] <- factor(r[,i])
  r
}

res_time2 <- mapply(function(n,m) {
  a <- experiment(createrandomdata(n,m),createrandomfactors(n),onlyy=TRUE)
  c(n,m,a["time1"],a["time2"])
  },
  rep(c(500,1000,5000,10000),25),
  rep(c(10,50,100,150,200),each=20))

saveRDS(res_time2,file="res_time2.rds")
```

Print the results.

```{r}
res_time2_median <- merge(aggregate(res_time2[3,],by=list(n=res_time2[1,],m=res_time2[2,]),median),
                          aggregate(res_time2[4,],by=list(n=res_time2[1,],m=res_time2[2,]),median),
                         by=c("n","m"))
res_time2_median <- res_time2_median[order(res_time2_median[,2]),]
for(i in 1:dim(res_time2_median)[1]) {
  cat(sprintf("$%d$ & $%d$ & $%.2f$ & $%.2f$ \\\\\n",
              res_time2_median[i,1],
              res_time2_median[i,2],
              res_time2_median[i,3],
              res_time2_median[i,4]))
}

```


# Exploration Part 1a: Exploration of the German data without background knowledge 

## Prepare tilings for exploration
Create the two hypothesis tilings for the exploration.
```{r}

## set the clustering threshold for choosng attributes using the parallel coordinates plots
thr <- 2/3

## Hypothesis 1 : all attributes are related
## Hypothesis 2 : no attributes are related
t0  <- tiling(nrow(data), ncol(data))
th1 <- t0$copy()
th2 <- t0$copy()

th1$addtile( C = seq.int(ncol(data)) )
```

## Visualise the data
The initial view of the data.

```{r}

pv  <- findproj2(th1$cov(data), th2$cov(data), k = 2)
u_vee <- pv$w
tmp_vee <- data %*% pv$w

data_h1 <- th1$permutedata(data)
data_h2 <- th2$permutedata(data)

```


## Step 1

We see a cluster in the upper right corner, which we now mark.

```{r}

ind_c1 <- c(333, 337, 338, 339, 341, 342, 344, 352, 353, 354, 355, 356, 357, 358, 359, 360, 362, 364, 365, 366, 367, 369, 370, 371, 372, 374, 375, 379, 380, 381, 382, 383, 384, 385, 386, 387, 388, 389, 397, 398, 399, 400, 401, 402, 403, 404, 405, 406, 407, 408, 409, 410, 411, 412)

plotandsave(data    = data,
            rdata1  = data_h1,
            rdata2  = data_h2,
            w       = pv$w,
            grouped = ind_c1,
            showsamples = FALSE,
            labs = TRUE,
	          ca = 1.25,
            fname = "view_ee.pdf",
            col_sel=sel_blue)

```

To see what features this cluster represents, we look at a parallel coordinates plot of the data.
                                                                                                    
```{r}

plotpcpandsave(data, ind_c1, fname = "pcplot_c1.pdf",col_sel=sel_blue)

```

To see which regions this corresponds to, we look at the original data which contains categorical variables describing the regions.

```{r}

tmp <- data_orig[ind_c1, ]
tab_c1 <- summary(tmp[, c("Region", "Type")])

print(tab_c1)
print(xtable(tab_c1), booktabs = TRUE, include.rownames = FALSE)

```    

It is clear that this cluster corresponds to rural districts in Eastern Germany and that these regions are characterised by a high degree of voting for the Left party.

We now add a tile for this observed pattern

```{r}

cluster_ci <- get_column_indices(data, ind_c1, thr)

th1$addtile( C = cluster_ci, R = ind_c1 )
th2$addtile( C = cluster_ci, R = ind_c1 )

data_h1 <- th1$permutedata(data)
data_h2 <- th2$permutedata(data)

```

## Step 2

The next step in the exploration.

```{r}

pv  <- findproj2(th1$cov(data), th2$cov(data), k = 2)
u_vet <- pv$w
tmp_vet <- data %*% pv$w

```


We now continue the exploration by turning our interest to the sparse cluster on the right side of the figure.

We mark the cluster.

We also see that the cluster we observed (the Eastern German cluster) is no longer prominent.

```{r}

ind_c2 <- c(1, 2, 3, 16, 17, 48, 49, 50, 63, 64, 65, 67, 71, 74, 80, 81, 92, 93, 99, 106, 107, 110, 118, 119, 120, 121, 137, 144, 155, 161, 164, 180, 194, 197, 205, 225, 285, 296, 326, 327, 328, 329, 330, 345, 346, 347, 348, 349, 350, 363, 368, 373, 376, 377, 378, 390, 391, 392, 393, 394)

plotandsave(data    = data,
         rdata1  = data_h1,
         rdata2  = data_h1[ind_c1,],
         w       = pv$w,
         grouped = ind_c2,
         showsamples = TRUE,
         labs = TRUE,
	 ca = 1.25,
         fname =  "view_et.pdf",
	       col_bg = sel_blue,
	       col_sel = sel_orange)


```

Again, to see which regions this corresponds to, we look at the original data which contains categorical variables describing the regions.

```{r}

plotpcpandsave(data, ind_c2, fname = "pcplot_c2.pdf",col_sel=sel_orange)

tmp <- data_orig[ind_c2, ]
tab_c2 <- summary(tmp[, c("Region", "Type")])

print(tab_c2)
print(xtable(tab_c2), booktabs = TRUE, include.rownames = FALSE)

```    
We conclude that this cluster represents urban regions.

We now add a tile for this observed pattern.

```{r}

cluster_ci <- get_column_indices(data, ind_c2, thr)

th1$addtile( C = cluster_ci, R = ind_c2 )
th2$addtile( C = cluster_ci, R = ind_c2 )

data_h1 <- th1$permutedata(data)
data_h2 <- th2$permutedata(data)

```
We conclude the exploration of the dataset.

# Exploration Part 1b: Exploration of the German data with a specific hypotheses 

Define the focus region

```{r}

ft_c_1 <- c("LEFT.2009", "CDU.2009", "SPD.2009", "FDP.2009", "GREEN.2009")
ft_c_2 <- c("Elderly.pop.", "Old.Pop.", "Mid.aged.Pop.", "Young.Pop.", "Children.Pop.")
ft_c_3 <- c("Agricult..workf.", "Prod..workf.", "Manufac..Workf.", "Constr..workf.", "Service.workf.", "Trade.workf.", "Finance.workf.", "Pub..serv..workf.")  
ft_c_4 <- c("Highschool.degree", "No.school.degree", "Unemploy.", "Unempl..Youth", "Income")

ft_ci_1 <- which( colnames(data) %in% ft_c_1)
ft_ci_2 <- which( colnames(data) %in% ft_c_2)
ft_ci_3 <- which( colnames(data) %in% ft_c_3)
ft_ci_4 <- which( colnames(data) %in% ft_c_4)

focus_tile_cind <- c(ft_ci_1, ft_ci_2, ft_ci_3, ft_ci_4)
focus_tile_rind <- which(data_orig$Type == "Rural")

t0  <- tiling(nrow(data), ncol(data))
th1 <- t0$copy()
th2 <- t0$copy()

th1$addtile( C = focus_tile_cind, R = focus_tile_rind)

for (tg in list(ft_ci_1, ft_ci_2, ft_ci_3, ft_ci_4)) {
    th2$addtile( C = tg, R = focus_tile_rind)
}

data_h1 <- th1$permutedata(data)
data_h2 <- th2$permutedata(data)

```

We now view the data.

```{r}

pv  <- findproj2(th1$cov(data), th2$cov(data), k = 2)
tmp <- data %*% pv$w

plotandsave(data    = data,
            rdata1  = data_h1,
            rdata2  = data_h2,
            w       = pv$w,
            grouped = focus_tile_rind,
            labs = TRUE,
            showsamples = FALSE,
            fname = "view_fe.pdf",
	          ca = 1.25,
            pfunc = plotdatafocus)

```

We observe a division into roughly two clusters. (Here we make some additional plots not included into the manuscript).

```{r}
ind_fc1 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 113, 114, 115, 116, 117, 118, 119, 120, 121, 122, 123, 124, 125, 126, 127, 128, 129, 130, 131, 132, 133, 134, 135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147, 148, 149, 150, 151, 152, 153, 154, 155, 156, 157, 158, 159, 160, 161, 162, 163, 164, 165, 166, 167, 168, 169, 170, 171, 172, 173, 174, 175, 176, 177, 178, 179, 180, 181, 182, 183, 184, 185, 186, 187, 188, 189, 190, 191, 192, 193, 194, 195, 196, 197, 198, 199, 200, 201, 202, 203, 204, 205, 206, 207, 208, 209, 210, 211, 212, 213, 214, 215, 216, 217, 218, 219, 220, 221, 222, 223, 224, 225, 226, 227, 228, 229, 230, 231, 232, 233, 234, 235, 236, 237, 238, 239, 240, 241, 242, 243, 244, 245, 246, 247, 248, 249, 250, 251, 252, 253, 254, 255, 256, 257, 258, 259, 260, 261, 262, 263, 264, 265, 266, 267, 268, 269, 270, 271, 272, 273, 274, 275, 276, 277, 278, 279, 280, 281, 282, 283, 284, 285, 286, 287, 288, 289, 290, 291, 292, 293, 294, 295, 296, 297, 298, 299, 300, 301, 302, 303, 304, 305, 306, 307, 308, 309, 310, 311, 312, 313, 314, 315, 316, 317, 318, 319, 320, 321, 323, 324, 325, 326, 368)

ind_fc2 <- c(322, 327, 328, 329, 330, 331, 332, 333, 334, 335, 336, 337, 338, 339, 340, 341, 342, 343, 344, 345, 346, 347, 348, 349, 350, 351, 352, 353, 354, 355, 356, 357, 358, 359, 360, 361, 362, 363, 364, 365, 366, 367, 369, 370, 371, 372, 373, 374, 375, 376, 377, 378, 379, 380, 381, 382, 383, 384, 385, 386, 387, 388, 389, 390, 391, 392, 393, 394, 395, 396, 397, 398, 399, 400, 401, 402, 403, 404, 405, 406, 407, 408, 409, 410, 411, 412)

ind_fc1 <- intersect(ind_fc1, focus_tile_rind)
ind_fc2 <- intersect(ind_fc2, focus_tile_rind)

```

We now visualise these two clusters in additional plots (not in the manuscript).

```{r}
plotandsave(data    = data,
            rdata1  = data_h1,
            rdata2  = data_h2,
            w       = pv$w,
            grouped = focus_tile_rind,
            cluster_ind = ind_fc1,
            fname = "view_fc1_large.pdf",
            showsamples = FALSE,
            showplot = FALSE,
            labs = TRUE,
            ca = 1.5,
            pfunc = plotdatafocus)

plotandsave(data    = data,
            rdata1  = data_h1,
            rdata2  = data_h2,
            w       = pv$w,
            grouped = focus_tile_rind,
            cluster_ind = ind_fc2,
            fname = "view_fc2_large.pdf",
            showsamples = FALSE,
            showplot = FALSE,
            labs = TRUE,
            ca = 1.5,
            pfunc = plotdatafocus)

```

```{r}
tab_fc1 <- summary(data_orig[ind_fc1, c("Region", "Type")])
tab_fc2 <- summary(data_orig[ind_fc2, c("Region", "Type")])

print(xtable(tab_fc1), booktabs = TRUE, include.rownames = FALSE)
print(xtable(tab_fc2), booktabs = TRUE, include.rownames = FALSE)

```

# Starting with some background knowledge

The initial view is different if we instead start with some background knowledge.

```{r}

## Add the tile corresponding to cluster 1 found in exploration
cluster_ci <- get_column_indices(data, ind_c1, thr)

th1$addtile(R = ind_c1, C = cluster_ci)
th2$addtile(R = ind_c1, C = cluster_ci)

data_h1 <- th1$permutedata(data)
data_h2 <- th2$permutedata(data)

```

In this case we would have found this view.

```{r}

pv  <- findproj2(th1$cov(data), th2$cov(data), k = 2)
tmp <- data %*% pv$w


plotandsave(data    = data,
            rdata1  = data_h1,
            rdata2  = data_h2,
            w       = pv$w,
            grouped = focus_tile_rind,
            showsamples = FALSE,
            labs = TRUE,
	    ca = 1.25,
            fname = "view_ft.pdf",
            pfunc = plotdatafocus)

```

# Calculating the gain for different projection vectors and hypothesis pairs

```{r}
## --------------------------------------------------
## Define a user tile spanning the first cluster
## found in exploration of the dataset
## --------------------------------------------------

tile_list_user <- list()
tile_list_user[[1]] <- list("rows" = ind_c1, "columns" = get_column_indices(data, ind_c1, thr))

## --------------------------------------------------
## Use the above defined focus tile
## --------------------------------------------------

tile_focus <- list()
tile_focus$column_groups <- list(ft_ci_1, ft_ci_2, ft_ci_3, ft_ci_4)
tile_focus$rows <- which(data_orig$Type == "Rural")

## --------------------------------------------------
## Calculate gains
## --------------------------------------------------

bold <- function(x) paste0("\\textbf{", x, "}")

tmp <- get_projection_vectors(data, tile_list_user, tile_focus)
res <- calculate_gains(data, tmp$u, tmp$cov_d)
res <- t(res)

print(res)

print(xtable(res, digits = 3), booktabs = TRUE, sanitize.colnames.function = bold, sanitize.rownames.function = bold)
```


# Exploration Part 2: Exploration of the accident data 

The code to run accident data exploration is commented, because we cannot distribute the proprietary data along the manuscript.

```{r}
# # Read in preprocessed data
# data <- readRDS("../webdemo/webdemo/data/srk_hdge.rds")
# 
# ## Hypothesis 1 : all attributes are related
# ## Hypothesis 2 : attributes inside each variable group are related, but there are no relations between the attribute groups
# t0  <- tiling(nrow(data), ncol(data))
# th1 <- t0$copy()
# th2 <- t0$copy()
# 
# # Let us create the attribute grouping
# # We drop variables "VALTIO" since it mostly  contains only "VALTIO.246"
# ci_list <- list(which(colnames(data) %in% c("AM1LK.0", "AM1LK.1", "AM1LK.2", "AM1LK.3", "AM1LK.4", "AM1LK.5", 
#                                  "AM1LK.6", "AM1LK.7", "AM1LK.8", "AM1LK.9", "AM1LK.A", "AM1LK.K","AM1LK.Y","AM1LK.L")), 
#      which( colnames(data) %in% c("EUSEUR.1", "EUSEUR.3", "EUSEUR.4", "EUSEUR.5", "EUSEUR.6", "EUSEUR.7", "EUSEUR.8", "EUSEUR.9")),
#      which( colnames(data) %in% c("IKAL.01", "IKAL.02", "IKAL.03", "IKAL.04", "IKAL.05", "IKAL.06", "IKAL.07", "IKAL.08", "IKAL.09",
#                                   "IKAL.10", "IKAL.11", "IKAL.12")),
#      which( colnames(data) %in% c("NUORET.1", "NUORET.2", "NUORET.3", "NUORET.4")),
#      which( colnames(data) %in% c("RUUMIS.11", "RUUMIS.12", "RUUMIS.13", "RUUMIS.14", "RUUMIS.15",  "RUUMIS.18",  "RUUMIS.19", 
#                                  "RUUMIS.21",  "RUUMIS.29", "RUUMIS.31",  "RUUMIS.39", "RUUMIS.41",  "RUUMIS.42",  "RUUMIS.43", 
#                                   "RUUMIS.48",  "RUUMIS.49",  "RUUMIS.51",  "RUUMIS.52",  "RUUMIS.53", "RUUMIS.54",  "RUUMIS.55", 
#                                   "RUUMIS.58",  "RUUMIS.61",  "RUUMIS.62",  "RUUMIS.63",  "RUUMIS.64",  "RUUMIS.65",  "RUUMIS.68", 
#                                   "RUUMIS.69","RUUMIS.71",  "RUUMIS.78", "RUUMIS.99", "RUUMIS.59")), #"RUUMIS.--"
#       which( colnames(data) %in% c("POIKKEA.10", "POIKKEA.20", "POIKKEA.30","POIKKEA.40", "POIKKEA.50", "POIKKEA.60", "POIKKEA.70",
#                                    "POIKKEA.80", "POIKKEA.99","POIKKEA.--")),
#       which( colnames(data) %in% c("SATK.1", "SATK.2", "SATK.3", "SATK.4", "SATK.5", "SATK.6", "SATK.7", "SATK.8", "SATK.9", "SATK.10",
#                                   "SATK.11", "SATK.12")),
#       which( colnames(data) %in% c("SUKUP.M", "SUKUP.N")),
#       which( colnames(data) %in% c("TOLP.A", "TOLP.B", "TOLP.C", "TOLP.D", "TOLP.E", "TOLP.F", "TOLP.G", "TOLP.H", "TOLP.I", 
#                                    "TOLP.J", "TOLP.K", "TOLP.L", "TOLP.M", "TOLP.N", "TOLP.O", "TOLP.P", "TOLP.Q", "TOLP.R", 
#                                    "TOLP.S", "TOLP.T", "TOLP.Z", "TOLP.--")),
#       which( colnames(data) %in% c("TPISTE.1", "TPISTE.2", "TPISTE.9", "TPISTE.--")), 
#       which( colnames(data) %in% c("TYOSUOR.10", "TYOSUOR.20", "TYOSUOR.30", "TYOSUOR.40", "TYOSUOR.50", "TYOSUOR.60", "TYOSUOR.70",
#                                     "TYOSUOR.99", "TYOSUOR.--")),
#       which( colnames(data) %in% c("TYOTEHT.11", "TYOTEHT.12", "TYOTEHT.19", "TYOTEHT.21", "TYOTEHT.22", "TYOTEHT.23", "TYOTEHT.24", 
#                                    "TYOTEHT.25", "TYOTEHT.29", "TYOTEHT.31", "TYOTEHT.32", "TYOTEHT.33", "TYOTEHT.34", "TYOTEHT.39",
#                                    "TYOTEHT.41", "TYOTEHT.42", "TYOTEHT.43", "TYOTEHT.49", "TYOTEHT.51", "TYOTEHT.52", "TYOTEHT.53",
#                                    "TYOTEHT.54", "TYOTEHT.55", "TYOTEHT.59", "TYOTEHT.61", "TYOTEHT.62", "TYOTEHT.63", "TYOTEHT.69",
#                                    "TYOTEHT.99","TYOTEHT.35", "TYOTEHT.--")),
#       which( colnames(data) %in% c("VAHITAP.11","VAHITAP.12", "VAHITAP.13", "VAHITAP.16", "VAHITAP.19", "VAHITAP.30", "VAHITAP.40", 
#                                    "VAHITAP.50", "VAHITAP.60", "VAHITAP.70", "VAHITAP.80", "VAHITAP.99", "VAHITAP.14", "VAHITAP.15",
#                                    "VAHITAP.17", "VAHITAP.--")), 
#       which( colnames(data) %in% c("VAHTY.M", "VAHTY.P")),
#       which( colnames(data) %in% c("VAMMAL.010", "VAMMAL.020", "VAMMAL.030", "VAMMAL.040", "VAMMAL.050", "VAMMAL.060", "VAMMAL.070", 
#                                    "VAMMAL.100", "VAMMAL.110", "VAMMAL.120", "VAMMAL.999","VAMMAL.090", "VAMMAL.--")), #"VAMMAL.--", 
#       which( colnames(data) %in% c("VPAIVA.1","VPAIVA.2","VPAIVA.3","VPAIVA.4","VPAIVA.5","VPAIVA.6", "VPAIVA.7")),
#       which( colnames(data) %in% c("VUOSI.2003", "VUOSI.2004", "VUOSI.2005", "VUOSI.2006", "VUOSI.2007", "VUOSI.2008", "VUOSI.2009", 
#                                    "VUOSI.2010", "VUOSI.2011", "VUOSI.2012", "VUOSI.2013", "VUOSI.2014")))
# # Add tile for Hypothesis 1
# focus_tile_cind <- sort(unlist(ci_list))
# th1$addtile(C = focus_tile_cind) # all rows
# 
# # Add tiles for Hypothesis 2
# for (tg in ci_list) {  
#     th2$addtile( C = tg)
# }
# 
# # Create samples for both the hypotheses
# data_h1 <- th1$permutedata(data)
# data_h2 <- th2$permutedata(data)
```


```{r}
# pv  <- findproj2(th1$cov(data), th2$cov(data), k = 2)
# 
# ind_c1_acc <- which(data %*% pv$w[,1] > 2.5e-5)
# 
# plotandsave(data    = data,
#             rdata1  = data_h1,
#             rdata2  = data_h2,
#             w       = pv$w,
#             grouped = ind_c1_acc,
#             showsamples = FALSE,
#             labs = TRUE,
# 	          ca = 1.25,
#             col_sel=sel_blue,
#             fname = "view_acc1.pdf")
```

To see which kind of accidents this selection corresponds to, we look at the original data and parallel coordinates plot. We see that these are accidents happening during travel to work ("VAHTY.M") and have also missing values (--) for  "TPISTE", "TYOTEHT", "TYOSUOR", "POIKKEA" and "VAHITAP" attributes (which are related to type of work/work place where the accident happened).

```{r}

# thr <- 2/3
# cluster_ci <- get_column_indices(data, ind_c1_acc, thr)
# 
# plotpcpandsave(data[,cluster_ci], ind_c1_acc, fname = "pcplot_acc1.pdf",col_sel=sel_blue)
# colnames(data[,cluster_ci])
# which(colMeans(data[ind_c1_acc,cluster_ci])>0)
# 
# compl_c1_acc <- setdiff(c(1:dim(data)[1]), ind_c1_acc)
# cluster_cci <- get_column_indices(data, compl_c1_acc, thr)
# which(colMeans(data[compl_c1_acc,cluster_cci])>0)

```   

Next we add a tile for the selection to both hypotheses, and compute a new projection. Now we observe that in this view the selection in blue is not more separating. Now one could continue the exploration further. 

```{r}
# th1$addtile( R = ind_c1_acc )
# th2$addtile( R = ind_c1_acc )
# 
# pv  <- findproj2(th1$cov(data), th2$cov(data), k = 2)
# 
# data_h1 <- th1$permutedata(data)
# data_h2 <- th2$permutedata(data)
# 
# plotandsave(data    = data,
#             rdata1  = data_h1,
#             rdata2  = data_h2,
#             w       = pv$w,
#             grouped = ind_c1_acc,
#             showsamples = FALSE,
#             labs = TRUE,
# 	          ca = 1.25,
#             col_sel=sel_blue,
#             fname = "view_acc2.pdf")

```


```{r}
# ind_c2_acc <- which(data %*% pv$w[,2] < -0.6e-4 & data %*% pv$w[,1] < -0.000015)
# 
# plotandsave(data    = data,
#             rdata1  = data_h1,
#             rdata2  = data_h1[ind_c1_acc,],
#             w       = pv$w,
#             grouped = ind_c2_acc,
#             showsamples = TRUE,
#             labs = TRUE,
# 	          ca = 1.25,
#             col_bg = sel_blue,
#             col_sel=sel_orange,
#             fname = "view_acc2.pdf")
# 
# 
# cluster_ci2 <- get_column_indices(data, ind_c2_acc, thr)
# plotpcpandsave(data[,cluster_ci2], ind_c2_acc, fname = "pcplot_acc2.pdf",col_sel=sel_orange)
# colnames(data[,cluster_ci2])
# which(colMeans(data[ind_c2_acc,cluster_ci2])>0)

```
